\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{pdfpages}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{blue},
    numberstyle=\tiny\color{codegray},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\title{Introduction to Artificial Intelligence Assignment \#1 report}
\author{Anton Brisilin, BS18-02 Student, \\ a.brisilin@innopolis.university}
\date{\today}
\begin{document}
\maketitle
    \section{Assumtions}
    The description of the Assignment is ambiguous, and require to make several
    assumtions in order to implement it:
    \begin{enumerate}
        \item Agent can not go out of bounds of field
        \item Cost of moving TO cell with human is 0, cost of moving FROM is 1
        \item For third algorithm, field is completely unknown and should be 
        explored to find the location of touchdown
        \item Agent ALWAYS present on cell (0,0), even if there is no \texttt{h(0,0)}
        in the input file
        \item If agent passes in wrong direction in random search, or in third algorithm
        search attempt fails (in case of third algorithm it means that whole search
        fails).
    \end{enumerate} 
    \section{Search algorithms description}
    \subsection{Random Search}
    The random search algorithm was implemented as a first one. At each turn 
    agent randomly perform either step or pass. There is almost nothing to say
    besides that there was eliminated possibility of agent to move to the 
    boundary of field and fail because of it.
    \subsection{Backtracking Search}
    The backtracking search tries all of possible paths and looking for the 
    shortest with DFS. However some optimisation I have added few optimisations 
    for it. First one is that backtracking search never visits the same node 
    twice - because shortest path to touchdown never contains loops. The second 
    optimisation is to keep track of already found paths and fail backtracking
    branches that are leading to path longer than one that is already found.
    \subsection{My own algorithm}
    The third implemented algorithm is not standart one. It is greedy algorithm 
    that on each step chooses the most unknown cell around agent, and then agent
    tries to go there. The algorithm is oriented to exploration, rather than 
    exploitation. The degree of cell's uncertainty is determined by the following
    function:\\ 
    $$\begin{cases}
        D(goal)=&0,\\
        D(ordinary\ cell) =& 1 - \frac{Visible-Unknown}{MaxVisible},\\
        D(cell\ with\ orc) =& 1\\
    \end{cases}$$
    In this function $Visible$ is number of cells, visible from the examined 
    one, $Unknown$ is number of cells, visible from examined that are unknown to
    the agent. $MaxVisible$ is maximal number of visible cells - it is depend 
    only on radius of vision. Agent always choose cell with least value to 
    examine next.  If the uncertainty degree of two or more adjacent cells are 
    equal, agent chooses one with cheaper movement cost (i.e. cell with human 
    will be preferred cell without one).\\
    On each step, algorithm keep $Reachable$ stack of cells that are not visited
    yet, but can be visited later. When all cells adjacent to agent are visited,
    $Reachable$ will be popped, and next cell to explore will be chosen from it.
    Path to this cell is searched with A* algorithm. If the $Reachable$ stack is
    empty, then algorithm fails.\\ 
    There is the limitation put in A*: it can visit only cells that were already 
    visited by the main algorithm. Other cells are considered as obstacles. 
    This limitation was put because A* operates in fully observable environment,
    whereas Assumption 3 states that the field is unknown. As a heuristics for 
    A* there was chosen manhattan distance to the goal.\\
    When the goal is reached, algorithm ends with success and path to the goal 
    point is returned.\\
    The algorithm is not capable of performing passes, because they does not make 
    sense in case of vision with radius 1 (there is no point in blind pass, because
    in this case search fails).\\
    \section{Testing algorithms on several maps}
    \subsection{Maps description}
    There were generated several maps of diferent types. For maps generation 
    there was written Python script, that produces file with Prolog predicates, 
    that can be given to Prolog interpreter as input knowledge base.\\
    Tested maps are listed below in text format: they are all 20*20 cells, with
    $(0,0)$ as the top-left cell, $X$ axis directed to right, $Y$ axis directed
    to the bottom of page.
    \subsubsection*{Legend}
    \begin{itemize}
        \item \texttt{O} - cell with orc
        \item \texttt{H} - cell with human
        \item \texttt{T} - cell with touchdown point
        \item \texttt{.} - empty cell
    \end{itemize}
    The following next pages contains considered maps together with time results
    and statistical analysis of time results.
    \newpage
    \subsubsection*{Map 1.}
    \lstinputlisting{maps/1}
    \textbf{Solvable} - YES.\\
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Search & Number of attempts & Best time & Mean time & Variance\\
            \hline
            Random Search & 100 $\times$ 100 & -- (no solutions found) & -- & --\\
            \hline
            Backtracking & 100 & 0.6194s & 0.5819s & 0.01134\\
            \hline
            My algorithm & 100 & 0.01466 & 0.0162 &  9.83e-6\\
            \hline
        \end{tabular}
    \end{center}
    \newpage
    \subsubsection*{Map 2.}
    \lstinputlisting{maps/2}
    \textbf{Solvable} - Only with passes.\\
    \textbf{Results:}.

    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Search & Number of attempts & Best time & Mean time & Variance\\
            \hline
            Random Search & 100 $\times$ 100 & -- (no solutions found) & -- & --\\
            \hline
            Backtracking & 100 & 0.0038s & 0.0043s & 7.3e-7\\
            \hline
            My algorithm & 100 & -- (no solutions found) & -- &  --\\
            \hline
        \end{tabular}
    \end{center}
    \newpage
    \subsubsection*{Map 3.}
    \lstinputlisting{maps/3}
    \textbf{Solvable} - Yes.\\
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Search & Number of attempts & Best time & Mean time & Variance\\
            \hline
            Random Search & 100 $\times$ 100 & -- (no solutions found) & -- & --\\
            \hline
            Backtracking & 100 & 0.0004s & 0.0006s & 9.18e-8\\
            \hline
            My algorithm & 100 & 0.1317 & 0.1405 &  0.00029\\
            \hline
        \end{tabular}
    \end{center}
    \newpage
    \subsubsection*{Map 4.}
    \lstinputlisting{maps/4}
    \textbf{Solvable} - Yes.\\
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Search & Number of attempts & Best time & Mean time & Variance\\
            \hline
            Random Search & 100 $\times$ 100 & 0.0025s & 0.0032s & 7.93e-7\\
            \hline
            Backtracking & 100 & 0.0552s & 0.0705s & 0.0003\\
            \hline
            My algorithm & 100 & 0.0002s & 0.0002s &  1.44e-8\\
            \hline
        \end{tabular}
    \end{center}
    \newpage
    \subsubsection*{Map 5.}
    \lstinputlisting{maps/5}
    \textbf{Solvable} - Yes.\\
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Search & Number of attempts & Best time & Mean time & Variance\\
            \hline
            Random Search & 100 $\times$ 100 & 0.0416s & 0.1407s & 0.0032\\
            \hline
            Backtracking & 100 & 0.07202s & 0.07648s & 4.1015e-6\\
            \hline
            My algorithm & 100 & 0.0003s & 0.0004s &  5.34e-8\\
            \hline
        \end{tabular}
    \end{center}
    \section{Algorithms comparison}
    Random search can not be considered as best algorithm to find paths. It is not
    reliable (not guarantee, that path will be found, if one exist). It has not 
    really big chances to find solution on orc-dense and/or human-sparse maps.
    I recommend to use Random search only in case when touchdown is really close 
    to start point, because in case of several lucky steps it can found best path
    to it faster than backtracking\\
    Backtracking search guaranties that path to the point will be found, also it
    can be used to obtain optimal path to the goal. But on orc-sparse maps, there
    are too many possible paths, so search for an optimal path can take unacceptably 
    long time (hours, days...). \\
    My algorithm works pretty well in case of both orc-sparse and orc-dense maps.
    Also it was designed in order to be utilised in partially observable environments.
    It works faster than even searching FIRST path with backtracking, in general case.
    However it did not found path on 2nd map, because, as was stated above, it 
    does not capable of doing passes, and 2nd map is unsolvable without passes.
    Also, in 3rd map my algorithm works slower, because of specific location of
    goal point: according to my heuristics, corner points has lowest priority to
    search in them.\\
    \section{Improved Vision}
    In second part of the assignment we was asked to compare same searching techniques  
    but with improved vision. However, for first and second algorithms it does 
    not really make sense, due to their nature: random search does not consider 
    vision at all, but backtracking, in contrast, works in fully observable environment.\\
    In case of third algorithm improved can make solvable maps, that can not be solved 
    without passes, because there is meaning in making passes now (when you see
    the person who you passing to).
    \section{Source code}
    Prolog source code files, together with Python maps generator are available
    on my Github repository, the link: \href{https://github.com/Mexator/IAI-Assignment}
    {https://github.com/Mexator/IAI-Assignment}.\\
    You can read \texttt{README.md} on the repository, to obtain instructions of 
    how to run the assignment.


\end{document}